---
title: "Prize Draw Selection"
author: "Robert Heirene"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: html
editor: visual
---

## Preamble

The purpose of this script is to randomly select 20 participants who completed a Qualtrics survey for our project with the overall title "Safer gambling online: Identifying and supporting at-risk consumers". These randomly selected participants will receive a voucher for their participation. The OSF page for this project can be found [here](https://osf.io/5at3q/).

```{r results = FALSE, message = FALSE}
#| code-summary: "Setup code"

library(tidyverse)
```

Read the data:

```{r results = FALSE, message = FALSE}
data  <-  read.csv("filepath") %>%
as_tibble()
```

Isolate those with contact details:

```{r results = FALSE, message = FALSE}
data_filtered  <-  data %>%
select(
Q4_1, # name
Q4_2, # email
Q4_3)  %>%# phone number
  mutate_all(~na_if(., "")) %>% # replace empty values with NA for easy filtering
  filter(!is.na(Q4_2) | !is.na(Q4_3)) # Only keep rows where a participant HAS included either their phone number or email address
```

Randomly select 20 people for the vouchers:

```{r results = FALSE, message = FALSE}
random_sample <- data_filtered %>%
  sample_n(20, replace = FALSE)


random_sample

write.csv(random_sample, "filepath")
```
